Re-run, Repeat, Reproduce, Reuse, Replicate:\\Transforming Code into Scientific Contributions

### 2017-10-16 draft.tex
Changes following review

Replicability is a cornerstone of science.  If an experimental result cannot be
re-obtained by an independent party, it merely becomes, at best, an observation
that may inspire future research \citep{Mesirov:2010,osc:2015}. Replication
issues have received increased attention in recent years, with a particular focus on medicine and psychology \citep{Iqbal:2016}.  One could think
that computer science would mostly be shielded from such issues, since a
computer program describes precisely what it does and is easily
disseminated to other researchers without alteration.

But precisely because it is easy to believe that if a program runs once and
gives the expected results it will do so forever, crucial steps to transform
working code into meaningful scientific contributions are rarely undertaken
\citep{Sandve:2013,Schwab:2000}. Computer science is plagued by replication
problems, in part, because it seems impervious to them. A program can
fail as a scientific contribution in many different ways for many different
reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
\citep{Goble:2016}, for a program to contribute to science, it should be
re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
reusable (R$^4$) and replicable (R$^5$). Let us illustrate this with a small
example, a random walk \citep{Hughes:1995} written in Python:\\

@@

% if code changes, parameters are lost
Recording the execution environment is only the first step. The R$^2$ program uses a random seed but does not keep a trace of it except in the code. Should the code change after the production of the results, someone provided with the last version of the code will not be able to know which seed was used to produce the results, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the results could obviously be found. For instance, seed 11235813 matches the results of R$^0$ with Python 2.7. This is not possible for a 100-step walk.}.\\

@@

Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}. Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user. This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper. With a documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\

@@

% reusable is not foolproof
Having made a software reusable offers an additional way to find errors, especially if your scientific contribution is popular. Unfortunately, this is not always effective, and some recent cases have dramatically illustrated the considerable impact a bug can have in science
\citep{Eklund:2016} or in our every-day life \citep{Durumeric:2014}. This is
why, as explained by Peng et al. \citep{Peng:2006}, {\em the replication of
  important findings by multiple independent investigators is fundamental to
  the accumulation of scientific evidence}.\\

% reproducible, a definition
\emph{Replicability} is the implicit
assumption that any article that does not provide the code source makes: that
the description it provides of the algorithms is sufficiently precise and
complete to re-obtain the results it presents. While every published article
should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\

@@

All these ambiguities in the description of an algorithm pile up; some are inconsequential (the 0 case has null probability), but some may affect the results in important ways. They are mostly inconspicuous to the reader and oftentimes, to the writer as well. In fact, the best way to ferret out the ambiguities, big and small, of an article is to replicate it. This is one of the reasons why the ReScience journal \citep{Rougier:2017} has been created. This journal {\em targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible}.\\

>>>>

Replicability is a cornerstone of science.  If an experimental result cannot be
re-obtained by an independent party, it merely becomes, at best, an observation
that may inspire future research \citep{Mesirov:2010,osc:2015}. Replication
issues have received increased attention in recent years, with a particular focus on medicine and psychology \citep{Iqbal:2016}.  One could think
that computational research would mostly be shielded from such issues, since a
computer program describes precisely what it does and is easily
disseminated to other researchers without alteration.

But precisely because it is easy to believe that if a program runs once and
gives the expected results it will do so forever, crucial steps to transform
working code into meaningful scientific contributions are rarely undertaken
\citep{Collberg:2016,Sandve:2013,Schwab:2000}. Computational research is plagued by replication
problems, in part, because it seems impervious to them. Contrary to production software who provide a service geared towards a practical outcome , the motivation behind scientific code is to test an hypothesis. While in some instance production software and scientific code are indistinguishable, the reason why they were created is different, and, therefore, so are the criteria to evaluate their success. A program can
fail as a scientific contribution in many different ways for many different
reasons. Borrowing the terms coined by \citeauthor{Goble:2016}
\citep{Goble:2016}, for a program to contribute to science, it should be
re-runnable (R$^1$), repeatable (R$^2$), reproducible (R$^3$),
reusable (R$^4$) and replicable (R$^5$). Let us illustrate this with a small
example, a random walk \citep{Hughes:1995} written in Python:\\

@@

% if code changes, parameters are lost
Recording the execution environment is only the first step. The R$^2$ program uses a random seed but does not keep a trace of it except in the code. Should the code change after the production of the results, someone provided with the last version of the code will not be able to know which seed was used to produce the results, and would need to iterate through all possible random seeds, an impossible task in practice\footnote{Here, with $2^{10}$ possibilities for a 10-step random walk, a seed matching the any possible sequence could certainly be found. For instance, seed 11235813 matches the results of R$^0$ with Python 2.7. Such a search becomes intractable for a 100-step walk.}.\\

@@

Similarly, code behavior should not be changed by commenting/uncommenting code \citep{Wilson:2017}. Modification of the behavior of the code, required when different experiments examine slightly different conditions, should always be explicitly set through parameters accessible to the end-user. This improves reproducibility in two ways: it allows those conditions to be recorded as parameters in the result files, and it allows to define separate scripts to run or configuration files to load to produce each of the figures of the published paper. With documentation explaining which script or configuration file corresponds to which experiment, reproducing the different figures becomes straightforward.\\

@@

% reusable is not foolproof
Having made a software reusable offers an additional way to find errors, especially if your scientific contribution is popular. Unfortunately, this is not always effective, and some recent cases have shown that bugs can lurk in well-used open-source code, impacting the false positive rates of fMRI studies \citep{Eklund:2016}, or the encryption of communications over the Internet \citep{Durumeric:2014}. Let's be clear: the goal here is not to remove all bugs and mistakes from science. The goal is to have methods and practices in place that make possible for the inevitable errors that will be made to be caught and corrected by motivated investigators. This is why, as explained by Peng et al. \citep{Peng:2006}, {\em the replication of
  important findings by multiple independent investigators is fundamental to
  the accumulation of scientific evidence}.\\

% reproducible, a definition
\emph{Replicability} is the implicit
assumption that any article that does not provide the source code makes: that
the description it provides of the algorithms is sufficiently precise and
complete to re-obtain the results it presents. Therefore, replicating a result means coding a new implementation from scratch that match the conceptual description of the article. This affords robustness to the results because, should the original code contain an error, a different codebase creates the possibility that this error will not be repeated, in the same way replicating an laboratory experiment can ferret out hidden biases. While every published article
should strive for replicability, it is seldom obtained. In fact, absent an explicit effort to make an algorithmic description replicable, there is little probability that it will be.\\

@@

All these ambiguities in the description of an algorithm pile up; some are inconsequential (the 0 case has null probability), but some may affect the results in important ways. They are mostly inconspicuous to the reader and oftentimes, to the writer as well. In fact, the best way to ferret out the ambiguities, big and small, of an article is to replicate it. This is one of the reasons why the ReScience journal \citep{Rougier:2017} was created (the second author, Nicolas Rougier, is one of the editor-in-chief of ReScience). This journal, run entirely by volunteers on GitHub, {\em targets computational research and encourages the explicit replication of already published research, promoting new and open-source implementations in order to ensure that the original research is reproducible}.\\
